---
title: "Relandscaping Cleaning Template"
format: html
editor: visual
---

## Relandscaping Data Cleaning: Carbon Flux

Winter 2024 - MCRS Advanced Quantitative Methods Class Project

Team #: 4

Team Members: Cherry, Janet, Katelin

### Load Packages

```{r}
library(tidyverse)
library(dplyr)
library(readxl)
```

### Import Data Files

```{r}
CFlux_Fall2018_raw <-read_excel("RelandscapingData/CFlux_Fall 2018 GH Relandscaping Carbon Flux Data.xlsx")
CFlux_June2018_raw <- read_excel("RelandscapingData/CFlux_GH Relandscaping Carbon Flux Data_Jun2018.xlsx")
CFlux_Winter2019_raw <- read_excel("RelandscapingData/CFlux_Winter 2019 GH Relandscaping Carbon Flux Data.xlsx")

```

### Create Primary

Reformat the data from each of your imported spreadsheets to have common columns and column formats so you can bind them into one object.

#### Data Frame #1: Fall 2018 data

```{r}
CFlux_Fall2018 <- CFlux_Fall2018_raw |>
  #remove the first row
  slice(-1)|>
  #change the "Date" column to "yyyy-mm-dd"
  mutate(Date = as.Date(as.numeric(Date), origin = "1899-12-30"))
```

#### Data Frame #2: June 2018 data

```{r}
CFlux_June2018 <- CFlux_Fall2018_raw |>
  #change 
  mutate()
```

#### Data Frame #3: Winter 2019 data

```{r}
CFlux_Winter2019 <- CFlux_Winter2019_raw |>
  #remove the first row
  slice(-1)|>
  #change the "Date" column to "yyyy-mm-dd"
  mutate(Date = as.Date(as.numeric(Date), origin = "1899-12-30"))
```

#### Complete Raw Data

Bind your cleaned data frames together into one object. Clean up any remaining column names or formats.

```{r}

```

### Cleaning

Go through your raw data frame column by column to find and correct issues.

#### Issue #1:

```{r}

```

#### Issue #2:

```{r}

```

etc.

```{r}

```

### Save your cleaned and complete Primary data frame to .csv and .Rdata and save the file as VARIABLE_Primary.csv or similar.

```{r}

```
