---
title: "Relandscaping Cleaning Template"
format: html
editor: visual
---

## Relandscaping Data Cleaning: Vacuum

Winter 2024 - MCRS Advanced Quantitative Methods Class Project

Team Members: Sully, Erin, Tom, Audrey

### Load Packages

```{r}
library('tidyverse')
library(readxl)

```

### Import Data Files

```{r}
Vacuum_RAW <- read_excel("RelandscapingData/Vacuum_‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏collect_data_arthropods_2019_7.12.21.xlsx", 
    sheet = "DATA", skip = 1, col_types="text")

```

### Create Primary

Reformat the data from each of your imported spreadsheets to have common columns and column formats so you can bind them into one object.

#### Data Frame #1: 

```{r}
# not applicable, no need to bind
```

#### Data Frame #2: fill in and add as needed

```{r}

# not applicable, omitting the metadata and just using the vacuum data
```

#### Complete Raw Data

Bind your cleaned data frames together into one object. Clean up any remaining column names or formats.

```{r}

# not applicable, only one dataframe
```

### Cleaning

Go through your raw data frame column by column to find and correct issues.

#### Issue #1: Changing dashes to underscores in sample name and column name

```{r}

Vacuum_Arthropods <- Vacuum_RAW |>
  mutate('Sample ID' = str_replace_all(`Sample ID`, "-", '_')) |> 
  rename(
    Sample_ID = `Sample ID`,
    Size_mm = `size (mm)`,
    count = `#`
  )
```

#### Issue #2: Remove unnecessary columns names

```{r}

Vacuum_Arthropods <- Vacuum_Arthropods |> 
  select(-c(photo, Date, `Date of sorting`, Tray, `ARCA code`, Cell, PlantRep, `Specific comments about specimen`, comments))
```

#### Issue #3: Convert numbers to numerics

```{r}

# convert count to numeric
Vacuum_Arthropods <- Vacuum_Arthropods |> 
  mutate(count = as.numeric(count))

# convert size to numeric
# 1. use "NA" for varies and na
Vacuum_Arthropods <- Vacuum_Arthropods |> 
  mutate(
    Is_Range = case_when(
      tolower(Size_mm) == 'varies' ~ NA, 
      tolower(Size_mm) == 'na' ~ NA,
      grepl("to", tolower(Size_mm), fixed = TRUE) ~ TRUE,
      .default = FALSE
    )
  ) |>
# 2. provide average for range values, and convert rest of numbers to numeric
  mutate (
    Size_mm = case_when(
      Size_mm == "3 to 5" ~ 4,
      Size_mm == "5 TO 10" ~ 7.5, 
      .default = as.numeric(Size_mm)
    )
  )
```

### Save your cleaned and complete Primary data frame to .csv and .Rdata and save the file as VARIABLE_Primary.csv or similar.

```{r}
write_csv(Vacuum_Arthropods, 'RelandscapingDataClean/Vacuum_Primary.csv')

save(Vacuum_Arthropods, file = 'RelandscapingDataClean/Vacuum_Primary.Rdata')
```
