---
title: "Relandscaping Cleaning Template"
format: html
editor: visual
---

## Relandscaping Data Cleaning: FILL IN YOUR VARIABLE HERE

Winter 2024 - MCRS Advanced Quantitative Methods Class Project

Team #:

Team Members:

### Load Packages

```{r}
library('tidyverse')
library(readxl)

```

### Import Data Files

```{r}
Vacuum_RAW <- read_excel("RelandscapingData/Vacuum_‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏‏collect_data_arthropods_2019_7.12.21.xlsx", 
    sheet = "DATA", skip = 1, col_types="text")

```

### Create Primary

Reformat the data from each of your imported spreadsheets to have common columns and column formats so you can bind them into one object.

#### Data Frame #1: Changing dashes to underscores in sample name

```{r}
Vacuum_Arthropods <- Vacuum_RAW |>
  mutate('Sample ID' = str_replace_all(`Sample ID`, "-", '_')) |> 
  select(-c(photo, Date, `Date of sorting`)) |> 
  rename(
    Sample_ID = `Sample ID`,
    Specimen_comments = `Specific comments about specimen`,
    ARCA_code = `ARCA code`,
    Size_mm = `size (mm)`
  )
```

#### Data Frame #2: fill in and add as needed

```{r}
```

#### Complete Raw Data

Bind your cleaned data frames together into one object. Clean up any remaining column names or formats.

```{r}
```

### Cleaning

Go through your raw data frame column by column to find and correct issues.

#### Issue #1:

```{r}


```

#### Issue #2:

```{r}

```

etc.

```{r}

```

### Save your cleaned and complete Primary data frame to .csv and .Rdata and save the file as VARIABLE_Primary.csv or similar.

```{r}

```
