---
title: "Relandscaping Cleaning Plant Size"
format: html
editor: visual
---

## Relandscaping Data Cleaning: FILL IN YOUR VARIABLE HERE

Winter 2024 - MCRS Advanced Quantitative Methods Class Project

Team #:

Team Members: Sully, Erin, Tom, Audrey

### Load Packages

```{r # install.packages(tidyverse)}
library(tidyverse)

```

### Import Data Files

```{r}
library(readxl)
columns <- c("Date", "Treatment","Block", "Observers", "Plant_Species", "Tag_Number", "Height_cm", "Width_cm", "Length_cm", "Notes"  )
Size_Fall_2018 <- read_excel("RelandscapingData/Size_Fall 2018 GH Relandscaping Plant Size Data.xlsx", 
    col_names = columns, skip = 2)

Size_Winter_2019 <- read_excel("RelandscapingData/Size_Winter 2019 GH Relandscaping Plant Size Data.xlsx",  col_names = columns, skip = 2)
```

### Create Primary

Reformat the data from each of your imported spreadsheets to have common columns and column formats so you can bind them into one object.

#### Data Frame #1: Fall 2018 data (example)

```{r}

```

#### Data Frame #2: fill in and add as needed

```{r}

```

#### Complete Raw Data

Bind your cleaned data frames together into one object. Clean up any remaining column names or formats.

```{r}
Size_2018_2019 <- bind_rows(Size_Fall_2018, Size_Winter_2019)
```

### Cleaning

Go through your raw data frame column by column to find and correct issues.

#### Issue #1:

```{r}
Size_2018_2019 = Size_2018_2019 |> 
  mutate("Plant_ID" = paste(Plant_Species, Tag_Number, sep='_')) |> 
  mutate(Date = ymd(Date)) |> 
  mutate(Year = year(Date))
```

#### Issue #2:

```{r}

```

etc.

```{r}

```

### Save your cleaned and complete Primary data frame to .csv and .Rdata and save the file as VARIABLE_Primary.csv or similar.

```{r}

```
