---
title: "Relandscaping Cleaning Template"
format: html
editor: visual
---

## Relandscaping Data Cleaning: Soil Moisture

Winter 2024 - MCRS Advanced Quantitative Methods Class Project

Team #: 3

Team Members:

### Load Packages

```{r}
library(tidyverse)
library(readr)
library(readxl)
library(lubridate)

```

### Import Data Files

```{r}

MoistS19 <- read_excel("RelandscapingData/Moist_ Spring 2019 GH Relandscaping Soil Moisture Data.xlsx", col_types = "c")
view(Moist19)
## No 2019 data

Moist18 <- read_excel("RelandscapingData/Moist_Fall 2018 GH Relandscaping Soil Moisture Data.xlsx", col_types = "text") %>% slice(-1)
view(Moist18)

Moist19<- read_excel("RelandscapingData/Moist_Winter 2019 GH Relandscaping Soil Moisture Data.xlsx", col_types ="text") %>% slice(-1)
view(Moist19)

```

### Create Primary

Reformat the data from each of your imported spreadsheets to have common columns and column formats so you can bind them into one object.

#### Complete Raw Data

Bind your cleaned data frames together into one object. Clean up any remaining column names or formats.

```{r}
Moist <- bind_rows(Moist18, Moist19)
```

### Cleaning

Go through your raw data frame column by column to find and correct issues.

#### Issue #1:

```{r}
```

#### Issue #2:

```{r}

```

etc.

```{r}

```

### Save your cleaned and complete Primary data frame to .csv and .Rdata and save the file as VARIABLE_Primary.csv or similar.

```{r}

```
