---
`---
title: "Relandscaping Cleaning Template"
format: html
editor: visual
---

## Relandscaping Data Cleaning: Carbon Flux

Winter 2024 - MCRS Advanced Quantitative Methods Class Project

Team #: 4

Team Members: Cherry, Janet, Katelin

### Load Packages

```{r}
library(tidyverse)
library(dplyr)
library(readxl)
```

### Import Data Files

```{r}
flux_fall18_raw <- read_excel("CFlux_Fall 2018 GH Relandscaping Carbon Flux Data.xlsx")

flux_jun18_raw <- read_excel("CFlux_GH Relandscaping Carbon Flux Data_Jun2018.xlsx")

flux_win19_raw <- read_excel("CFlux_Winter 2019 GH Relandscaping Carbon Flux Data.xlsx")


```

### Create Primary

Reformat the data from each of your imported spreadsheets to have common columns and column formats so you can bind them into one object.

#### Data Frame #1: Fall 2018 data

```{r}
#### Data Frame #1: Fall 2018 data (example)

flux_fall18_date <- flux_fall18_raw |>
  #remove first row
  slice(-1) |>
  #change date in "Date" column to "yyyy-mm-dd"
  mutate(Date = as.Date(as.numeric(Date), origin = "1899-12-30", ))

```

#### Data Frame #2: Winter 2019 data

```{r}
flux_win19_date <- flux_win19_raw |>
  #remove first row
  slice(-1) |>
  #change date in "Date" column to "yyyy-mm-dd"
  mutate(Date = as.Date(as.numeric(Date), origin = "1899-12-30", ))
```

**Data Frame #3: June 2018 data**

```{r}
flux_jun18_date <-flux_jun18_raw |>
  #change last column to state "Notes"
  rename("Notes" = "...9") #change last column to state "Notes"

```

#### Complete Raw Data

Bind your cleaned data frames together into one object. Clean up any remaining column names or formats.

```{r}
flux_fall18_ready <- flux_fall18_date |> #time as.numeric, CO2 as.numeric
  mutate(`Time (sec)` = as.numeric(`Time (sec)`)) |>
  mutate(`CO2 (ppm)` = as.numeric(`CO2 (ppm)`))

flux_win19_ready <- flux_win19_date |>   #time as.numeric, CO2 as.numeric
  mutate(`Time (sec)` = as.numeric(`Time (sec)`)) |>
  mutate(`CO2 (ppm)` = as.numeric(`CO2 (ppm)`))

flux_jun18_ready <- flux_jun18_date |>   #date as.date
  mutate(Date = as.Date(Date))

fall18win19 <- bind_rows(flux_fall18_ready, flux_win19_ready)

flux_data_comb <- bind_rows(fall18win19, flux_jun18_ready)
```

### Cleaning

Go through your raw data frame column by column to find and correct issues.

Cherry branch: cherry_phase1

Katelin branch: katelin_phase1

Janet branch: janet_phase1

**Cleaning:**

-   **reformat names (no spaces, "-" or "\_" only)**

-   **variables in appropriate format**

-   **mostly free of data errors**

-   **primary file saved as both .csv and .Rdata**

-   [**ANNOTATED**]{.underline}

#### Issue #1:

```{r}

```

#### Issue #2:

```{r}

```

etc.

```{r}

```

### Save your cleaned and complete Primary data frame to .csv and .Rdata and save the file as VARIABLE_Primary.csv or similar.

```{r}

```
