---
title: "Relandscaping Cleaning Template"
format: html
editor: visual
---

## Relandscaping Data Cleaning: FILL IN YOUR VARIABLE HERE

Winter 2024 - MCRS Advanced Quantitative Methods Class Project

Team #: 2

Team Members:

### Load Packages

```{r}
#install.packages(tidyverse)
library(tidyverse)
library(ggplot2)
library(readr)
library(readxl)

```

### Import Data Files

```{r}
Quad_Winter_19 <- read_excel("Quad_Winter 2019 GH Relandscaping Quadrat Data.xlsx", 
             sheet = 1)
            # col_names = FALSE, 
             # col_names = TRUE,
             #col_types = TRUE, 
            #col_types = "text" 

```

```{r}
Quad_Spring_18 <- read_excel("Quad_GH Relandscaping Quadrat Data_Jun2018.xlsx")
```

```{r}
QuadFall_18<- read_excel("Quad_Fall 2018 GH Relandscaping Quadrat Data.xlsx")
```

### Create Primary

Reformat the data from each of your imported spreadsheets to have common columns and column formats so you can bind them into one object.

#### Data Frame #1: Fall 2018 data (example)

```{r}

```

#### Data Frame #2: fill in and add as needed

```{r}

```

#### Complete Raw Data

Bind your cleaned data frames together into one object. Clean up any remaining column names or formats.

```{r}

```

### Cleaning

Go through your raw data frame column by column to find and correct issues.

#### Issue #1:

```{r}

```

#### Issue #2:

```{r}

```

etc.

```{r}

```

### Save your cleaned and complete Primary data frame to .csv and .Rdata and save the file as VARIABLE_Primary.csv or similar.

```{r}

```
