---
title: "Soil Moisture Relandscaping Cleaning"
format: html
editor: visual
---

## Relandscaping Data Cleaning: Soil Moisture

Winter 2024 - MCRS Advanced Quantitative Methods Class Project

Team #: 3

Team Members:Tory, Jazz, and Kelsey

### Load Packages

```{r}
library(tidyverse)
library(readr)
library(readxl)
library(lubridate)

```

### Import Data Files

```{r}

MoistS19 <- read_excel("RelandscapingData/Moist_ Spring 2019 GH Relandscaping Soil Moisture Data.xlsx")
#view(MoistS19)
## No 2019 data
#Import and remove row without data
Moist18 <- read_excel("RelandscapingData/Moist_Fall 2018 GH Relandscaping Soil Moisture Data.xlsx", col_types = "text") %>% slice(-1)
#Replace values with NA that were entered into the wrong collumn
Moist18[Moist18 == "soil"] <- NA
Moist18[Moist18 == "tough"] <- NA
#view(Moist18)
#import and remove collumns without data
MoistW19<- read_excel("RelandscapingData/Moist_Winter 2019 GH Relandscaping Soil Moisture Data.xlsx", col_types ="text") %>% slice(-1,-4)
#view(MoistW19)

```

### Create Primary

Reformat the data from each of your imported spreadsheets to have common columns and column formats so you can bind them into one object.

#### Complete Raw Data

Bind your cleaned data frames together into one object. Clean up any remaining column names or formats.

```{r}
Moist <- bind_rows(Moist18, MoistW19)
View(Moist)
```

### Cleaning

Go through your raw data frame column by column to find and correct issues.

#### Issue #1: correct date since its weird

```{r}
Moist1 <-Moist|>
  mutate(Date=
as.Date(as.numeric(Date), origin = "1899-12-30"))
view(Moist1)
```

### Save your cleaned and complete Primary data frame to .csv and .Rdata and save the file as VARIABLE_Primary.csv or similar.

```{r}
# Kelsey 
# turns data into a CSV file, we just replace Moist1 with whatever our final data set is named 

write.csv(Moist1, "RelandscapingData/MOIST_Primary.csv", row.names = FALSE)
save(Moist1, file="RelandscapingData/MOIST_Primary.RData")
```

```{r}
#Jazz EDA

```

```{r}
#Tory EDA
#load color
library(ggplot2)
#Question: Do the species have different Soil Moisture VWC(volumetric water content)?
#adjust Canopy1_VWC so its numeric
moist2<-transform(Moist1,Canopy1_VWC=as.numeric(Canopy1_VWC))
sapply(moist2, mode)

#COLORS
customcolors <- c("#556B2F","#FFB90F")

#plot as boxplot
plotCanVWC.Species <-ggplot(moist2, 
  aes( x =Plant.Species,y =Canopy1_VWC, fill=Treatment))+
  geom_boxplot()

p<-plotCanVWC.Species + scale_fill_manual(values = customcolors)
print(p+labs(y="Soil Moisure(VWC)",x="Species"))

```

```{r}
#Kelsey EDA
# how does the VWC for btwn plots differ between blocks
  
Moist1$'Btwn1_VWC' <- as.numeric(Moist1$'Btwn1_VWC')
Moist1$'Btwn2_VWC' <- as.numeric(Moist1$'Btwn2_VWC')


#making a new column for water consumption with the averqge of the Btwn1 and Btwn2 values 
Moist1$Btwn_VWC = ((Moist1$Btwn1_VWC + Moist1$Btwn2_VWC)/2)


MoistPlot <- ggplot(Moist1, aes(x = `Plant Species`, y = Btwn_VWC, fill = Treatment)) + 
    geom_boxplot() +
  facet_grid(.~Date)


MoistPlot
  #geom_bar(stat = "identity") +
  #ggtitle("Bar Plot Example") +
  #ylab("Values") +
  #xlab("Categories")
```
