---
title: "Transect Reland Data Cleaning"
format: html
editor: visual
---

## Relandscaping Data Cleaning: Soil Moisture

Winter 2024 - MCRS Advanced Quantitative Methods Class Project

Team #: 3

Team Members:

### Load Packages

```{r}
library(tidyverse)
library(readr)
library(readxl)
library(lubridate)

```

### Import Data Files

```{r}

# Transect data

TransectF18 <- read_excel("RelandscapingData/Transect_VegHeight_Fall 2018 GH Relandscaping Transect Data.xlsx", sheet = 2) |> 
  slice(-1)
  
view(TransectF18) 

TransectS18 <- read_excel("RelandscapingData/Transect_VegHeight_GH Relandscaping Transect Data_Jun2018.xlsx", sheet = 2 )
  
view(TransectS18)

TransectW19 <- read_excel("RelandscapingData/Transect_Winter 2019 GH Relandscaping Transect Data.xlsx") |> 
  slice(-1)
  
view(TransectW19) 





```

### Create Primary

Reformat the data from each of your imported spreadsheets to have common columns and column formats so you can bind them into one object.

#### Complete Raw Data

Bind your cleaned data frames together into one object. Clean up any remaining column names or formats.

```{r}
Transect <- bind_rows(TransectF18, TransectS18, TransectW19)
View(Transect)
```

### Cleaning

Go through your raw data frame column by column to find and correct issues.

#### Issue #1: correct date since its weird

```{r}
# fixing the date

TransectW19 <-TransectW19|>
  mutate(Date=
as.Date(as.numeric(Date), origin = "1899-12-30"))

TransectF18 <-TransectF18|>
  mutate(Date=
as.Date(as.numeric(Date), origin = "1899-12-30"))

```

#### Issue #2:

```{r}
# S18 siteID is a num needs to be a character
TransectS18$'Site ID' <- as.character(TransectS18$'Site ID')

glimpse(TransectS18)

## S18 distance is a num needs to be a character

TransectS18$'Distance' <- as.character(TransectS18$'Distance')

glimpse(TransectW19)

```

etc.

```{r}

```

### Save your cleaned and complete Primary data frame to .csv and .Rdata and save the file as VARIABLE_Primary.csv or similar.

```{r}

```
